Clustering Beat-Chroma Patterns in a Large Music Database

### 2010-03-23 Paper/Paper.tex
improved, removed rate distortion section
cowbell data description added

\subsection{Training: Cowbell Dataset}\label{sec:traindata}
We have $43,300$ songs, giving  
%$\bf{3,720,091}$ 
$\mathbf{3.7}$\textbf{M} 
non silent bars, a bar usually 
contains $4$ beats. WHERE DOES THIS DATA COMES FROM? GIFT FROM EN?
\footnote{http://www.morecowbell.dj/}

@@

\subsection{Music Randomness}\label{ssec:random}
It will not surprise any reader, music is not random. This fact should
become apparent in our experiments in the following way. Information
theory tells us that if the signal to encode doubles, we need to square
the size of the codebook to encode it with the same distortion.

@@

% see generate_expo_codes_curves.py to generate the following figure:
\begin{figure}[htb]
\begin{center}
\includegraphics[width=.99\columnwidth]{codesize_patternsize}
\end{center}
\caption{\small{
Experiment on squaring the codebook size when the pattern size doubles.
Each line represent $4$ codebook sizes: $2$, $4$, $16$ and $256$.
The pattern size range from $1$ for $1/4$ of a bar to $32$ for $8$ bars.
If everything was random, each of these lines should be flat. The training
size was fixed to $128$K samples.
}}
\label{fig:size_pattern}
\end{figure}

@@

\subsection{Cluster}
Once we understand the kind of codebook we create, we naturally want
to visualize the clusters, \ie the patterns from actual song that
surround a given centroid. Unsurprisingly, real patterns tend to be
noisier than codes. However, centroids seem to capture important aspects
of the patterns. We could compare the centroids with the result of
a lowpass filtering.

See Figure \ref{fig:cluster} and \ref{fig:cluster_diff}.

>>>>

\subsection{Training: Cowbell Dataset}\label{sec:traindata}
We have $43,300$ tracks that were uploaded
to morecowbell.dj\footnote{http://www.morecowbell.dj/}, 
an online service that uses the Echo
Nest to produce a remixed version of
an uploaded music file that adds cowbell and other sound
effects in time with the music.
The $43.3$K songs, giving $\mathbf{3.7}$\textbf{M}
non silent bars, a bar usually
contains $4$ beats (Figure \ref{fig:barsize}).

@@

% REMOVING THE FOLLOWING SUBSECTION ON RATE - DISTORTION
\iffalse
\subsection{Music Randomness}\label{ssec:random}
It will not surprise any reader, music is not random. This fact should
become apparent in our experiments in the following way. Information
theory tells us that if the signal to encode doubles, we need to square
the size of the codebook to encode it with the same distortion.

@@

% see generate_expo_codes_curves.py to generate the following figure:
\begin{figure}[htb]
\begin{center}
\includegraphics[width=.99\columnwidth]{codesize_patternsize}
\end{center}
\caption{\small{
Experiment on squaring the codebook size when the pattern size doubles.
Each line represent $4$ codebook sizes: $2$, $4$, $16$ and $256$.
The pattern size range from $1$ for $1/4$ of a bar to $32$ for $8$ bars.
If everything was random, each of these lines should be flat. The training
size was fixed to $128$K samples.
}}
\label{fig:size_pattern}
\end{figure}
\fi

@@

\subsection{Cluster}
Once we understand the kind of codebook we create, we naturally want
to visualize the clusters, \ie the patterns from actual song that
surround a given centroid. Unsurprisingly, real patterns tend to be
noisier than codes. However, centroids seem to capture important aspects
of the patterns. We could compare the centroids with the result of
a lowpass filtering. In Figure \ref{fig:cluster} we show elements of the same
cluster, and in Figure \ref{fig:cluster_diff}, we emphasize the difference
with the centroid.
